\bf Classification and Regression on Random Dot Product Graphs

### 2024-11-15 draft.tex
first draft of paper

\bigskip
\begin{abstract}
The random dot product graph (RDPG) has become a powerful modeling tool
in uncovering latent structures within graphs. In particular, it has
been shown that the RDPG describes a wide range of popular random graph
models with rigid latent structures. More recently, joint modeling of
mutliple random graphs that share common properties or structures across
graphs have been introduced, such as the multilayer RDPG, multiple RPDG,
and multilayer stochastic block model. In this work, we use these joint
random graph models in the context of statistical learning, such as
classification and regression, by introducing the multiple latent
structure model, in which the graphs share a common latent structure
with different parameters that correspond to different response
variables. Then we propose various estimation techniques involving
manifold learning to estimate these parameters and in turn predict the
responses, with theorems guaranteeing convergence of the predictions.
Simulations, as well as applications on brain connectivity networks,
verify the performance of our methods.
\end{abstract}

@@

Graph and network data are now as ubiquitous as traditional feature data
in the fields of sociology (e.g., social networks), neuroimaging (e.g.,
brain connectivity networks), and deep learning (e.g., graph neural
networks). As a result, new statistical and machine learning methods
have recently been developed to analyze network data. One approach is to
treat the network as a random graph that comes from some probability
model. In particular, if we sum up the network in an adjacency matrix
\(A \in \mathbb{R}^{n \times n}\), then one probability model might be
to draw each element \(A_{ij}\), which represents the existence of an
edge or the edge weight from vertex \(i\) to \(j\), independently from
some distribution, perhaps with a unique parameter for the pair
\((i,j)\), e.g., \(A_{ij} \indep F_{\theta_{ij}}\). The classical
example of this is the \(\ER\) model \citep{Gilbert:1959}, in which
every edge is drawn from the same distribution, typically a Bernoulli
distribution, using the same parameter, i.e.,
\(A_{ij} \iid \Bernoulli (p)\). The inhomogeneous Bernoulli graph
extends this by allowing each edge, represented by \(A_{ij}\), to have
its own parameter, \(P_{ij}\), e.g., in the Bernoulli case,
\(A_{ij} \indep \Bernoulli(P_{ij})\). Typically, the parameters are
collected into an edge probability matrix (in the case of unweighted
graphs) or edge parameter matrix (in the case of weighted graphs),
denoted as \(P \in \mathbb{R}^{n \times n}\). The network analysis
problem in this setting is to estimate \(P\) given \(A\).

If the parameter matrix \(P\) is unconstrained, the inference problem is
overparameterized. On the other hand, the classical \(\ER\) model is
often too restrictive to describe real, observed networks. Much work has
been done to develop models that are constrained enough for robust
statistical inference while being generalizable to describe a wide range
of networks. One such family of random graph models is the random dot
product graph (RDPG), first proposed by
\citet{10.1007/978-3-540-77004-6_11}, which is a type of latent space
graph in which each vertex of the graph has a corresponding latent
vector in a low-dimensional Euclidean space \(\mathbb{R}^d\), and the
edge parameter between each pair of vertices is determined by the dot
product of the corresponding vectors. In this model, the constraint is
the low rank of the parameter matrix \(P\), assuming that the latent
dimension \(d\) is less than the number of vertices \(n\). Further
constraints can be imposed on the RDPG in the form of distributional
assumptions on the latent vectors or restricting the latent vectors to
lie on subspaces or manifolds in the latent space
\citep{athreya2020estimation}.

@@

In this paper, we apply these highly structured GRDPG models to
supervised learning problems in which there are multiple graphs,
\(G_1, ..., G_L\), each with a response \(y^{(1)}, ..., y^{(L)}\). Each
graph, represented by adjacency matrix \(A^{(l)}\), is presumed to be
drawn from a GRDPG with parameter set \(\theta_l\), and in turn, the
response \(y_l\) is drawn from a univariate distribution with parameter
\(\theta_l\). Since the latent vectors of each \(A^{(l)}\) are presumed
to be distributed on a structured subset of the latent space, e.g., a
manifold, we call this the multiple latent structure model (MLSM). Much
previous work has been done on multilayer networks, such as the
multilayer degree corrected block model of \citet{agterberg2022joint},
multiple random eigen graphs of \citet{8889404}, and the multiple common
subspace independent edge graphs of \citet{arroyo2020inference}. In each
of these models, the main inference problem is to identify parameters
that are shared across multiple graphs. Our contribution is in applying
similar methods to identify parameters specific to each graph, which are
in turn used as features for supervised learning.

The following sections are organized as follows: Section 2 introduces
the structured and parameterized

\section{Methods}\label{methods}

@@

Let \(G_1, ..., G_L\) be a collection of undirected graphs where each
\(G_l = (V_l, E_l)\). In this setup, each graph can have its own vertex
and edge sets. Each graph is represented by adjacency matrices
\(A^{(1)}, ..., A^{(L)}\), where each
\(A^{(l)} \in \mathbb{R}^{n_l \times n_l}\) is a symmetric and hollow
matrix such that \(A^{(l)}_{ij}\) represents the edge weight between
vertices \(i\) and \(j\), and \(n_l\) represents the size of the vertex
set of the \(l\)\textsuperscript{th} graph. Let
\(P^{(1)}, ..., P^{(L)}\) be the corresponding collection of edge
parameter matrices such that each
\(A^{(l)}_{ij} \indep H(P^{(l)}_{ij})\) for each
\(1 \leq i < j \leq n_l\) and \(H\) is a specified probability
distribution such that \(E[A^{(l)}_{ij}] = P^{(l)}_{ij}\). For ease of
notation, we use \(A^{(l)} \sim H(P^{(l)})\) for an adjacency matrix
\(A^{(l)}\) drawn from edge parameter matrix \(P^{(l)}\) in this way.
Denote \(X^{(1)}, ..., X^{(L)}\) as the sets of latent vectors
corresponding to each graph under the GRDPG framework. Each
\(X^{(l)} \in \mathbb{R}^{n_l \times d}\), i.e., the graphs share the
same latent dimension, and \(x_i^{(l)} \in \mathbb{R}^d\) is the latent
vector corresponding to the \(i\)\textsuperscript{th} vertex in the
\(l\)\textsuperscript{th} graph.

@@

We propose our methodology for classification and regression in three
steps:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Curve-fitting algorithm to recover the LSM curves from the ASEs.
\item
  Quasi-maximum likelihood estimation to recover the parameters of the
  underlying distributions.
\item
  Classification or regression to determine the relationship between the
  parameters and the responses.
\end{enumerate}

@@

For curve-fitting, we will restrict the LSM curves to
non-self-intersecting Bezier polynomials \citep{10.5555/320367}. A
polynomial curve of degree \(R\) in \(\mathbb{R}^d\) is uniquely defined
by a Bezier polynomial within the space of Bezier polynomials of degree
at most \(R\), up to reverse order, although there are infinitely many
Bezier polynomials of degree greater than \(R\) that can describe the
curve \citep{SANCHEZREYES2022102118}. We further restrict the curves to
the case \(p(0) = 0\), i.e., the curve begins at the origin. This is
consistent with well-known models such as the DCBM and PABM when viewed
as LSMs or mixtures of LSMs
\citep{Koo_2022, rubindelanchy2017statistical}, as well as real networks
that can adequately be described as LSMs \citep{athreya2020estimation}.
This constraint also removes the reverse order nonidentifiability. Thus,
if \(R\), the degree of the polynomial, can be determined, and the curve
begins at the origin, a collection of \(n \geq R\) unique vectors on the
curve and not at the origin can determine the unique Bezier polynomial
that describes the curve.

@@

For initialization, the ordering of \(t_1, ..., t_n\) is determined via
a one-dimensional Isomap embedding \citep{Tenenbaum2000-ff}.

@@

\begin{proof}

\end{proof}

\subsubsection{Recovery of the Underlying
Distributions}\label{recovery-of-the-underlying-distributions}

\begin{algorithm}[H]
\label{alg:bezier-param-fit}
\DontPrintSemicolon
\SetAlgoLined
\setstretch{1.1}
\KwData{Adjacency matrix $A$, embedding dimension $d$, probability distribution family $F$, polynomial degree $R$, Isomap neighborhood parameter $\lambda$, stopping criterion $\epsilon$.}
\KwResult{Bezier coefficients $\hat{b}$, timepoint values $\hat{t}_1, ..., \hat{t}_n$.}
Compute $\hat{X} \in \mathbb{R}^{n \times d}$, the $d$-dimensional ASE of $A$.\;
Initialize timepoints $\hat{t}_1, ..., \hat{t}_n$, first via a one-dimensional Isomap embedding of $\hat{X}$ using neighborhood parameter $\lambda$, then by normalizing the embedding to $[0, 1]$.\;
\Repeat {the change in equation \ref{eq:bezier-ls} is less than $\epsilon$.} {
  Fit $\hat{b}$ by equation \ref{eq:bezier-ols}.\;
  \For {$i = 1, ..., n$} {
    Fit $\hat{t}_i$ by equation by finding the roots of equation \ref{eq:dbezier} and choosing the roots that minimizes equation \ref{eq:bezier-ls}.\;
  }
}
Estimate $\hat{\theta}$ from $\hat{t}_1, ..., \hat{t}_n$ via maximum likelihood estimation.\; 
\caption{Procedure for estimating the underlying distribution of a Bezier LSM curve from an adjacency matrix.}
\end{algorithm}

@@

\begin{algorithm}[H]
\label{alg:mlsm-fit}
\DontPrintSemicolon
\SetAlgoLined
\setstretch{1.1}
\KwData{Adjacency matrices $A^{(1)}, ..., A^{(L)}$, response variables $y_1, ..., y_L$, embedding dimension $d$, probability distribution family $F$, polynomial degree $R$, Isomap neighborhood parameters $\lambda_1, ..., \lambda_L$, stopping criterion $\epsilon$.}
\KwResult{Bezier coefficients $\hat{b}$, timepoint values $\hat{t}_1, ..., \hat{t}_n$.}
\For {$\ell = \{1, ..., L\}$} {
  Compute $\hat{X}^{(\ell)} \in \mathbb{R}^{n_\ell \times d}$, the $d$-dimensional ASE of $A^{(\ell)} \in \mathbb{R}^{n_\ell \times n_\ell}$.\;
  Initialize timepoints $\hat{t}_1, ..., \hat{t}_{n_\ell}$, first via a one-dimensional Isomap embedding of $\hat{X}^{(\ell)}$ using neighborhood parameter $\lambda_\ell$, then by normalizing the embedding to $[0, 1]$.\;
  \Repeat {the change in equation \ref{eq:bezier-ls} is less than $\epsilon$.} {
    Fit $\hat{b}$ by equation \ref{eq:bezier-ols}.\;
    \For {$i = 1, ..., {n_\ell}$} {
      Fit $\hat{t}_i$ by equation by finding the roots of equation \ref{eq:dbezier} and choosing the roots that minimizes equation \ref{eq:bezier-ls}.\;
    }
  }
  Estimate $\hat{\theta}_\ell$ from $\hat{t}_1, ..., \hat{t}_{n_\ell}$ via maximum likelihood estimation.\;
}
Fit a model for predicting $y_1, ..., y_L$ using the estimated parameters $\theta_1, ..., \theta_L$ (e.g., via linear regression).\;
\caption{Procedure for fitting a classification or regression model for an MLSM.}
\end{algorithm}

\begin{theorem}
\ref{thm:supervised}
\end{theorem}

@@

In this section, we evaluate algorithm \ref{alg:mlsm-fit} and theorem
\ref{thm:supervised} empirically via simulation and compare to existing
methods. Two settings are examined. In the first, the data generating
model reduces to a simpler model and can be adequately described by the
COSIE model or multilayer DCBM. The second setting involves a nonlinear
latent structure that can be described as a polynomial curve and thus
does not reduce to a simpler model or be analyzed by their corresponding
methods.

The scenarios are as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  In this setting, there are two latent structures, described by two
  rays emanating from the origin.
\item
\end{enumerate}

\subsection{Regression}\label{regression}

In this scenario, the responses \(y_\ell\) are generated as linear
responses of \(\theta_\ell\) (in scenario 1) or \(a_\ell\) and
\(b_\ell\) (in scenario 2). More specifically, in scenario 1, each
\(y_\ell \indep \mathcal{N}(\beta_0 + \beta_1 \theta_\ell, \sigma^2)\),
and in scenario 2, each
\(y_\ell \indep \mathcal{N}(\beta_0 + \beta_1 a_\ell + \beta_2 b_\ell, \sigma^2)\).

@@

\subsection{Classification}\label{classification}

In this simulation experiment, the latent vectors were sampled along a
Bezier curve defined by \(g(t) = \bigl[ t^2 \; 2 t (1-t) \bigr]^\top\).
The timepoints \(t_i\) were sampled as iid Beta random variables with
two sets of parameters, \((\alpha_1, \beta_1)\) and
\((\alpha_2, \beta_2)\). The setup is as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Draw response variables \(y_1, ..., y_n \iid \Multinomial(1/2, 1/2)\).
\item
  For each \(i = 1, ..., n\), draw
  \(t_i \mid y_i \indep \Betadist(\alpha_{y_i}, \beta_{y_i})\), such
  that \(\alpha_1 = 1\), \(\beta_1 = 2\), \(\alpha_2 = 2\), and
  \(\beta_2 = 1\).
\item
  Construct each latent vector as \(x_i = g(t_i)\) and compile them in
  data matrix \(X = \bigl[ x_1 \; \cdots \; x_n \bigr]^\top\).
\item
  Sample graph and its adjacency matrix as \(A \sim \RDPG(X)\).
\end{enumerate}

For each graph, we constructed the ASE, which was used to estimate the
parameters \((\hat{\alpha}, \hat{\beta})\) for the graph, using the
maximum likelihood method. The estimated parameters were then used as
predictors \(y_1, ..., y_n\), setting aside half for training and half
for testing. We investigated graphs of size \(|V| = 32, 64, 128, 256\).
The number of graphs for each experiment was set to \(L = 64\). For each
(number of graphs, size of graph) pair, we performed 32 replicates.
Figure \ref{fig:class-sim-ase} shows the ASE of one graph.\\
Figure \ref{fig:classification-sim} shows the boxplots of the
classification error rates.

{\centering \includegraphics{draft_files/figure-latex/class-sim-ase-1}

\caption{One simulated graph (left) and its ASE (right). The red curve is the fitted quadratic Bezier curve on the ASE.}\label{fig:class-sim-ase}
\end{figure}

{\centering \includegraphics{draft_files/figure-latex/classification-sim-1}

\caption{Median classification error rate and its IQR vs. number of vertices in each graph.}\label{fig:classification-sim}
\end{figure}

\subsection{Regression}\label{regression-1}

In the first regression simulation,

In the next regression simulation, we applied

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Draw angles
  \(\theta_1, ..., \theta_N \iid \Uniform(\pi / 6, \pi / 3)\).
\item
  For each \(k = 1, ..., N\),

  \begin{enumerate}
  \def\labelenumii{\roman{enumii}.}
  \tightlist
  \item
    Draw \(t_1, ..., t_n \iid \Uniform(0, 1)\);
  \item
    Draw \(z_1, ..., z_n \iid \Multinomial(1/2, 1/2)\);
  \item
    For each \(i = 1, ..., n\), set
    \(x_i = \begin{cases} \bigl[ t_i \; 0 \bigr]^\top \; z_i = 1 \\ \bigl[ t_i \cos \theta_\ell \; t_i \sin \theta_\ell \bigr]^\top \; z_i = 2 \end{cases}\);
  \item
    Collect \(X = \bigl[ x_1 \; \cdots \; x_n \bigr]^\top\) and draw
    \(A \sim \RDPG(X)\);
  \item
    Set the response \(y_\ell = \beta_0 + \beta_1 \theta_\ell\)
  \end{enumerate}
\end{enumerate}

In this simulation, we set \(\beta_0 = \beta_1 = 1\). The number of
graphs was set to \(L = 64\), and the number of vertices per graph was
set to \(n = 128, 256, 512, 1024\). For each \(n\), we simulated 32
replicates.

\begin{center}\includegraphics{draft_files/figure-latex/angle-reg-example-1} \end{center}

\begin{center}\includegraphics{draft_files/figure-latex/angle-reg-results-1} \end{center}

@@

\caption{Graphs of the Drosophila connectomes. The left and right are of the left and right hemispheres, respectively. Each vertex represents a neuron, which are labeled by neuron type.}\label{fig:mbconnectome-graph}
\end{figure}

@@

\caption{ASEs of the Drosophila connectome graphs. The embedding vectors are labeled by neuron type. }\label{fig:mbconnectome-ase}
\end{figure}

>>>>

\bigskip
\begin{abstract}
The random dot product graph (RDPG) has become a powerful modeling tool
in uncovering latent structures within graphs. In particular, it has
been shown that the RDPG describes a wide range of popular random graph
models with rigid latent structures. More recently, joint modeling of
mutliple random graphs that share common properties or structures across
graphs have been introduced, such as the multilayer RDPG, multiple RPDG,
and common subspace independent edge model. In this work, we use these
joint random graph models in the context of graph classification and
regression by introducing the multiple latent structure model, in which
the graphs share a common latent structure with different parameters
that correspond to different response variables. Then we propose various
estimation techniques involving manifold learning to estimate these
parameters and in turn predict the responses, with theorems guaranteeing
convergence of the predictions. Simulations, as well as applications on
brain connectivity networks, verify the performance of our methods.
\end{abstract}

@@

Graph and network data are now as ubiquitous as traditional feature data
in the fields of sociology (e.g., social networks), neuroimaging (e.g.,
brain connectivity networks), and deep learning (e.g., graph neural
networks). As a result, new statistical and machine learning methods
have recently been developed to analyze network data. One approach is to
treat the network as a random graph that comes from some probability
model. One such approach in particular is the independent edge approach.
If we sum up the network in an adjacency matrix
\(A \in \mathbb{R}^{n \times n}\), then this model draws each element
\(A_{ij}\), which represents the existence of an edge or the edge weight
from vertex \(i\) to \(j\), independently from some distribution,
perhaps with a unique parameter for the pair \((i,j)\), e.g.,
\(A_{ij} \indep F_{\theta_{ij}}\). The classical example of this is the
\(\ER\) model \citep{Gilbert:1959}, in which every edge is drawn from
the same distribution, typically a Bernoulli distribution, using the
same parameter, i.e., \(A_{ij} \iid \Bernoulli (p)\). The inhomogeneous
Bernoulli graph extends this by allowing each edge, represented by
\(A_{ij}\), to have its own parameter, \(P_{ij}\), e.g., in the
Bernoulli case, \(A_{ij} \indep \Bernoulli(P_{ij})\). Typically, the
parameters are collected into an edge probability matrix (in the case of
unweighted graphs) or edge parameter matrix (in the case of weighted
graphs), denoted as \(P \in \mathbb{R}^{n \times n}\). The network
analysis problem in this setting is to estimate \(P\) given \(A\).

If the parameter matrix \(P\) is unconstrained, the inference problem is
overparameterized. On the other hand, the classical \(\ER\) model is
often too restrictive to describe real, observed networks. Various
approaches have been taken to restrict \(P\), such as by setting the
matrix rank of \(P\) to a small value. One such family of random graph
models is the random dot product graph (RDPG), first proposed by
\citet{10.1007/978-3-540-77004-6_11}, which is a type of latent space
graph in which each vertex of the graph has a corresponding latent
vector in a low-dimensional Euclidean space \(\mathbb{R}^d\), and the
edge parameter between each pair of vertices is determined by the dot
product of the corresponding vectors. In this model, the constraint is
the low rank of the parameter matrix \(P\), assuming that the latent
dimension \(d\) is less than the number of vertices \(n\). Further
constraints can be imposed on the RDPG in the form of distributional
assumptions on the latent vectors or restricting the latent vectors to
lie on subspaces or manifolds in the latent space
\citep{athreya2020estimation}.

@@

In this paper, we apply these highly structured random dot product graph
models to supervised learning problems in which there are multiple
graphs, \(G_1, ..., G_L\), each with a response
\(y^{(1)}, ..., y^{(L)}\). Each graph, represented by adjacency matrix
\(A^{(\ell)}\), is presumed to be drawn from a RDPG with parameter set
\(\theta_l\), and in turn, the response \(y_l\) is drawn from a
univariate distribution with parameter \(\theta_l\). Since the latent
vectors of each \(A^{(\ell)}\) are presumed to be distributed on a
structured subset of the latent space, e.g., a manifold, we call this
the multiple latent structure model (MLSM). Much previous work has been
done on multilayer networks, such as the multilayer degree corrected
block model of \citet{agterberg2022joint}, multiple random eigen graphs
of \citet{8889404}, and the common subspace independent edge graphs of
\citet{arroyo2020inference}. In each of these models, the main inference
problem is to identify parameters that are shared across multiple
graphs. Our contribution is in applying similar methods to identify
parameters specific to each graph, which are in turn used as features
for supervised learning. This further allows us to model graphs with
differing vertex sets.

The remainder of this paper is organized as follows: Section 2 builds on
the latent structure model to introduce the multiple latent structure
model and proposes manifold learning algorithms for estimation. Sections
3 and 4 provide simulation studies and real data analyses. Section 5
concludes. Proofs of theorems are provided in the Appendix.

\section{Setup and Methodology}\label{setup-and-methodology}

@@

Let \(G^{(1)}, ..., G^{(L)}\) be a collection of undirected graphs where
each \(G^{(\ell)} = (V^{(\ell)}, E^{(\ell)})\). In this setup, each
graph can have its own vertex and edge sets. These graphs are
represented by adjacency matrices \(A^{(1)}, ..., A^{(L)}\), where each
\(A^{(\ell)} \in \mathbb{R}^{n_l \times n_l}\) is a symmetric and hollow
matrix such that \(A^{(\ell)}_{ij}\) represents the edge weight between
vertices \(i\) and \(j\), and \(n_l\) represents the size of the vertex
set of the \(l\)\textsuperscript{th} graph. Let
\(P^{(1)}, ..., P^{(L)}\) be the corresponding collection of edge
parameter matrices such that each
\(A^{(\ell)}_{ij} \indep H(P^{(\ell)}_{ij})\) for each
\(1 \leq i < j \leq n_l\) and \(H\) is a specified probability
distribution such that \(E[A^{(\ell)}_{ij}] = P^{(\ell)}_{ij}\). For
ease of notation, we use \(A^{(\ell)} \sim H(P^{(\ell)})\) for an
adjacency matrix \(A^{(\ell)}\) drawn from edge parameter matrix
\(P^{(\ell)}\) in this way. Denote \(X^{(1)}, ..., X^{(L)}\) as the sets
of latent vectors corresponding to each graph under the GRDPG framework.
Each \(X^{(\ell)} \in \mathbb{R}^{n_\ell \times d}\), i.e., the graphs
share the same latent dimension, and \(x_i^{(\ell)} \in \mathbb{R}^d\)
is the latent vector corresponding to the \(i\)\textsuperscript{th}
vertex in the \(\ell\)\textsuperscript{th} graph. \(\hat{X}^{(\ell)}\)
is the adjacency spectral embedding of \(A^{(\ell)}\) and is an
estimator for \(X^{(\ell)}\).

@@

We propose our methodology for classification and regression in three
steps: First, we propose a curve-fitting algorithm to recover the LSM
curves from the ASEs. Second, we the parameters of the underlying
distributions from the curve fit via maximum likelihood estimation.
Third, we apply classification or regression methods to determine the
relationship between the estimated parameters and the responses.
Finally, we provide theorems that show the consistency of the curve fits
and maximum likelihood estimators.

@@

For curve-fitting, we will restrict the LSM curves to
non-self-intersecting Bezier polynomials \citep{10.5555/320367}. A
polynomial curve of degree \(R\) defined by
\(p : [0, 1] \to \mathbb{R}^d\) is uniquely defined by a Bezier
polynomial within the space of Bezier polynomials of degree at most
\(R\), up to reverse order, although there are infinitely many Bezier
polynomials of degree greater than \(R\) that can describe the curve
\citep{SANCHEZREYES2022102118}. We further restrict the curves to the
case \(p(0) = 0\), i.e., the curve begins at the origin. This is
consistent with well-known models such as the DCBM and PABM when viewed
as LSMs or mixtures of LSMs
\citep{Koo_2022, rubindelanchy2017statistical}, as well as real networks
that can adequately be described as LSMs \citep{athreya2020estimation}.
This constraint also removes the reverse order nonidentifiability. Thus,
if \(R\), the degree of the polynomial, can be determined, and the curve
begins at the origin, a collection of \(n \geq R\) unique vectors on the
curve and not at the origin can determine the unique Bezier polynomial
that describes the curve.

@@

@@

\subsubsection{\texorpdfstring{Quasi-Maximum Likelihood Estimation for
\(\theta\)}{Quasi-Maximum Likelihood Estimation for \textbackslash theta}}\label{quasi-maximum-likelihood-estimation-for-theta}

The estimation of \(T\) and the restrictions imposed on the family of
Bezier curves allows for the estimation of \(\theta\) via maximum
likelihood estimation (algorithrm \ref{alg:bezier-param-fit}). Since
\(\hat{T}\) is a consistent estimator for \(T\) by theorem
\ref{thm:bezier-consistency}, if \(t_1, ..., t_n \sim F_\theta\), a
maximum likelihood estimator using \(\hat{t}_1, ..., \hat{t}_n\) is a
consistent estimator for \(\theta\). Algorithm
\ref{alg:bezier-param-fit} details the procedure.

\begin{algorithm}[H]
\label{alg:bezier-param-fit}
\DontPrintSemicolon
\SetAlgoLined
\setstretch{1.1}
\KwData{Adjacency matrix $A$, embedding dimension $d$, probability distribution family $F$, polynomial degree $R$.}
\KwResult{Bezier coefficients $\hat{b}$, timepoint values $\hat{t}_1, ..., \hat{t}_n$.}
Compute $\hat{X} \in \mathbb{R}^{n \times d}$, the $d$-dimensional ASE of $A$.\;
Initialize timepoints $\hat{t}_1, ..., \hat{t}_n$.\;
\Repeat {the change in equation \ref{eq:bezier-ls} is less than $\epsilon$.} {
  Fit $\hat{b}$ by equation \ref{eq:bezier-ols}.\;
  \For {$i = 1, ..., n$} {
    Fit $\hat{t}_i$ by equation by finding the roots of equation \ref{eq:dbezier} and choosing the roots that minimizes equation \ref{eq:bezier-ls}.\;
  }
}
Estimate $\hat{\theta}$ from $\hat{t}_1, ..., \hat{t}_n$ via maximum likelihood estimation.\; 
\caption{Procedure for estimating the underlying distribution of a Bezier LSM curve from an adjacency matrix.}
\end{algorithm}

@@

Applying algorithm \ref{alg:bezier-param-fit} to supervised learning
problems is straightforward. First, we extract each
\(\hat{\theta}^{(\ell)}\) from each graph, then apply a supervised
learning algorithm (e.g., ordinary least squares) for predicting each
\(y^{(\ell)}\) from each \(\hat{\theta}^{(\ell)}\). Since
\ref{alg:bezier-param-fit} produces consistent estimators, for large
graphs the supervised learning algorithm in this regime will perform
just as well as if we already knew each \(\theta^{(\ell)}\). Therefore,
\(\tilde{\beta}\), the estimator based on
\(\hat{\theta}^{(1)}, ..., \hat{\theta}^{(L)}\), is consistent for
\(\beta\) (theorem \ref{thm:supervised}).

\begin{algorithm}[H]
\label{alg:mlsm-fit}
\DontPrintSemicolon
\SetAlgoLined
\setstretch{1.1}
\KwData{Adjacency matrices $A^{(1)}, ..., A^{(L)}$, response variables $y_1, ..., y_L$, embedding dimension $d$, probability distribution family $F$, polynomial degree $R$, stopping criterion $\epsilon$.}
\KwResult{Bezier coefficients $\hat{b}$, timepoint values $\hat{t}_1, ..., \hat{t}_n$.}
\For {$\ell = \{1, ..., L\}$} {
  Compute $\hat{X}^{(\ell)} \in \mathbb{R}^{n_\ell \times d}$, the $d$-dimensional ASE of $A^{(\ell)} \in \mathbb{R}^{n_\ell \times n_\ell}$.\;
  Initialize timepoints $\hat{t}_1, ..., \hat{t}_{n_\ell}$;
  \Repeat {the change in equation \ref{eq:bezier-ls} is less than $\epsilon$.} {
    Fit $\hat{b}$ by equation \ref{eq:bezier-ols}.\;
    \For {$i = 1, ..., {n_\ell}$} {
      Fit $\hat{t}_i$ by equation by finding the roots of equation \ref{eq:dbezier} and choosing the roots that minimizes equation \ref{eq:bezier-ls}.\;
    }
  }
  Estimate $\hat{\theta}^{(\ell)}$ from $\hat{t}_1, ..., \hat{t}_{n_\ell}$ via maximum likelihood estimation.\;
}
Fit a model for predicting $y^{(\ell)}, ..., y^{(\ell)}$ using the estimated parameters $\theta^{(\ell)}, ..., \theta^{(\ell)}$ (e.g., via linear regression).\;
\caption{Procedure for fitting a classification or regression model for an MLSM.}
\end{algorithm}

\begin{theorem}
\label{thm:supervised} 
Suppose $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{p^{(\ell)}\}, F, \{\theta^{(\ell)}\})$ where $p : [0, 1] \to \mathbb{R}^d$ is a Bezier polynomial of degree $R < d$ and $p(0) = 0$. 
Let each $y^{(\ell)} \indep \mathcal{N}\big( (\theta^{(\ell)})^\top \beta, \sigma^2 \big)$, and let $\tilde{\beta}$ be the ordinary least squares estimate of $\beta$ using $\tilde{\theta}^{(1)}, ..., \tilde{\theta}^{(L)}$, the outputs of \ref{alg:bezier-param-fit} for each graph. 
Then as each $n_\ell \to \infty$, $\tilde{\beta} \stackrel{p}{\to} \beta$. 
\end{theorem}

@@

In this section, we evaluate algorithm \ref{alg:mlsm-fit} and theorem
\ref{thm:supervised} empirically via simulation and compare to existing
methods. Two settings are examined. In the first, the data generating
model reduces to a simpler model and can be adequately described by the
COSIE model or multilayer DCBM. The second setting involves graphs of
variable sizes drawn from a nonlinear latent structure that can be
described as a polynomial curve and thus does not reduce to a simpler
model or be analyzed by their corresponding methods.

In these simulations, the responses \(y^{(\ell)}\) are generated as
linear responses of \(\theta^{(\ell)}\) (in scenario 1) or
\(a^{(\ell)}\) and \(b^{(\ell)}\) (in scenario 2). More specifically, in
scenario 1, each
\(y_\ell \indep \mathcal{N}(\beta_0 + \beta_1 \theta^{(\ell)}, \sigma^2)\),
where \(\theta^{(\ell)}\) is the angle between two line segments in the
latent space, and in scenario 2, each
\(y_\ell \indep \mathcal{N}(\beta_0 + \beta_1 a^{(\ell)} + \beta_2 b^{(\ell)}, \sigma^2)\),
where \(a^{(\ell)}\) and \(b^{(\ell)}\) describe the Beta parameters of
the distribution of latent vectors along a curve. We set
\(\sigma^2 = 0.25\) for both scenarios.

The simulation setup for the first scenario is as follows: For each
\(N = 128, 256, 512, 1024\), sample \(L = 128\) graphs as:

\begin{enumerate}
\def\labelenumi{\roman{enumi}.}
\tightlist
\item
  Draw \(n_1, ..., n_L\) as \(n_\ell \iid \Poisson(N)\).
\item
  Draw
  \(\theta^{(1)}, ..., \theta^{(L)} \iid \Uniform(\pi / 6, \pi / 3)\).
\item
  For each \(\ell = 1, ..., L\), draw latent vectors either from the
  line segment described by
  \(p_1(t) = \begin{bmatrix} t & 0 \end{bmatrix}^\top\) or
  \(p_2(t) = \begin{bmatrix} \cos(\theta^{(\ell)}) & \sin(\theta^{(\ell)}) \end{bmatrix}^\top\)
  with probability \(1/2\) for each.
\item
  Use the \(L\) sets of latent vectors to draw \(L\) RDPGs with
  adjacency matrices \(A^{(1)}, ..., A^{(L)}\).
\item
  Draw each
  \(y_\ell \indep \mathcal{N}(1 + 2 \theta^{(\ell)}, \sigma^2)\).
\item
  Apply algorithm \ref{alg:mlsm-fit} to construct a linear model
  connecting \(y_1, ..., y_\ell\) and \(A^{(1)}, ..., A^{(L)}\), using
  the first half as training data and reporting MSE on the last half not
  used for training.
\end{enumerate}

For the second the simulation setup is similar. For each
\(N = 128, 256, 512, 1024\), sample \(L = 128\) graphs as:

\begin{enumerate}
\def\labelenumi{\roman{enumi}.}
\tightlist
\item
  Draw \(n_1, ..., n_L\) as \(n_\ell \iid \Poisson(N)\).
\item
  Draw each
  \((a^{(\ell)}, b^{(\ell)}) \iid \Uniform([0.5, 2] \times [0.5, 2])\).
\item
  For each \(\ell = 1, ..., L\), draw latent vectors by first taking
  \(t_1, ..., t_{n_\ell} \sim \Betadist(a^{(\ell)}, b^{(\ell)})\) then
  taking each latent vector as
  \(x_i = \begin{bmatrix} t^2 & 2 t (1-t) \end{bmatrix}^\top\).
\item
  Use the \(L\) sets of latent vectors to draw \(L\) RDPGs with
  adjacency matrices \(A^{(1)}, ..., A^{(L)}\).
\item
  Draw each
  \(y_\ell \indep \mathcal{N}(1 + 2 \alpha^{(\ell)} + b^{(\ell)}, \sigma^2)\).
\item
  Apply algorithm \ref{alg:mlsm-fit} to construct a linear model
  connecting \(y_1, ..., y_\ell\) and \(A^{(1)}, ..., A^{(L)}\), using
  the first half as training data and reporting MSE on the last half not
  used for training.
\end{enumerate}

In figure \ref{fig:angle-regression-sim}, we compare algorithm
\ref{alg:mlsm-fit} to fitting a regression model using graph
transitivity and average degree of each vertex. Since this scenario is
similar to a multilayer DCBM, we expect performance to be very similar,
which is what we observe in the results. However, we cannot model this
simulation as a multilayer DCBM since although each graph can be
described as a DCBM, the graphs do not share the same vertex set. In
figure \ref{fig:beta-regression-sim}, we see that linear regression
using graph summary statistics performs better for smaller graphs, but
as we draw more vertices, the ASE improves and so algorithm
\ref{mlsm-fit} performs better, whereas graph summary statistics cannot
improve with the number of vertices. Figures \ref{fig:theta-mse} and
\ref{fig:beta-mse} show that the quasi-MLE estimators computed via
algorithm \ref{alg:bezier-param-fit} are consistent.

@@

@@

\caption{Graphs of the Drosophila connectomes. The left and right are of the left and right hemispheres, respectively. Each vertex represents a neuron, which are labeled by neuron type. The red vertices are Kenyon Cells, the blue vertices are Input and Output Neurons, and the green vertices are Projection Neurons.}\label{fig:mbconnectome-graph}
\end{figure}

@@

\caption{ASEs of the Drosophila connectome graphs. The embedding vectors are labeled by neuron type, with red being Kenyon Cells, blue being Input and Output Neurons, and green being Projection Neurons.}\label{fig:mbconnectome-ase}
\end{figure}
